---
title: "ETC5521 Assignment 1"
subtitle: "Global Astronautical exploration trends"
team: quoll
author:
  - Jinhao Luo
  - Yawen Zhang
date: "`r Sys.Date()`"
bibliography: references.bib
biblio-style: apa
output: 
  bookdown::html_document2:
    citation_package: biblatex
    toc: true
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, message=FALSE, warning=FALSE)

library(tidyverse)
library(dplyr)
library(plotly)
library(ggplot2)
```

```{r write-data, eval = FALSE}
tuesdata <- tidytuesdayR::tt_load('2020-07-14')
astronauts_origin <- tuesdata$astronauts

write.csv(astronauts_origin, here::here("data", "astronauts.csv"))
```

```{r read-data}
astronauts <- read.csv(here::here("data", "astronauts.csv"))
```

# Introduction and motivation

@corlett2020population indicated that since the first humans completed an orbital spaceflight in 1961 and stepped on the moon in 1969, space missions have been completed nearly 60 years. Meanwhile, space exploration is still technologically challenging and competitive. Therefore, this report would try to analyses the development of the world astronautic exploration from 1961. In the following sections, we will describe the data set and source of the data, and what is the data prepared for. The data description also includes how we transform and clean the raw data for analysis. Beside, our statistical programming used for analysis is R and Rstudio. However, the data set might be not comprehensive, which might result in insufficient analysis and conclusion. Meanwhile, some missing values also included in the data, which might impact on the accuracy of the report. Such limitation should be considered and improved in following analysis.

Based on the purpose of finding the development of astronautical exploration, this report would consider a primary question and three secondary questions. The primary question will focus on the development of the world astronautic exploration with time from the first spaceflight launched. In addition, the three secondary questions should be: Firstly, When does astronautic exploration become more competitive? Secondly, How is the tech of astronautic exploration development? Thirdly,  Do females also play an important role in astronautic exploration, when does the first female astronaut appear in international astronautic exploration?

# Data description

The data which are used for analysing in this report are from the astronaut database. Because only one website could not contain comprehensive and up-to-date astronautic information. Therefore, this database was collected from NASA, Roscosmos, and fun-made websites, which includes information about the space mission before 15 January 2020. Furthermore, this data contains 24 different types of information about 1277 astronauts which are name, sex, year of birth, nationality, military status, the name and year of selection program, information about each mission implemented by astronaut such as year, title, ascend and decent shuttle names, duration of mission and extravehicular activity (EVAs). The version of data has been updated on 14 July 2020, which contains the astronautic information from 1961 to 2019. The data used in this report was read from the TidyTuesdayR package, and it has been saved in a new csv file called astronaut for later use.

By browsing the whole dataset and using the glimpse function to observe the variables and the corresponding classes by R, this report decides to apply this dataset into analysing the development of the astronautic industry. Meanwhile, according to the observation, the report would try to use several important and useful variables to explore the development of the astronautic industry as much as possible, like the year, number, and duration of missions, proportion of space travelers, the number of extravehicular activity and the gender of astronauts. In addition, some missing values and inappropriate classes of variables have been found, which would be considered in the following analysing part.

# Analysis and findings

To overview the global astronautical exploration trends, we first look at the changes of the astronautical exploration projects since the first human being to explore the outer space. Furthermore, human’s enthusiasm for exploring the space is also reflected on the tech of astronautic exploration development. Last but not least, although men as the main force in spaceflights, female also played an important role in recent years. The following sections will concentrate on three secondary questions to explore the global astronautical exploration trends. 

## Trends in the astronautical exploration

@corlett2020population have indicated that humans have never stop their steps of space exploration. Since 1961 the first humans have accomplished an orbital spaceflight, many countries join the team of exploring the space. Therefore, identify the trend in astronautical exploration projects could help the readers realise the development of the space exploration. According to figure \@ref(fig:yearly-total-projects), the general trend of global space exploration projects is increase from 1961 to 1985. In 1985, the number of global project has reached the peak which 28 projects has been launched in this single year. After that the global project has experienced a significant decrease trend. Although the number of project has bounced to 14 in 1992, and fluctuated in the following seventeen years, the general trend of project amount is decrease after 1985. In general, based on figure \@ref(fig:yearly-total-projects) which might indicates that the global astronautical exploration competition between each government is fierce, and the competition has reached the peak in 1985. Although the amount has decreased after peak. The competition, however, still remained strong until 2009.

```{r yearly-total-projects, fig.cap= "Yearly astronauticalal exploration project amount in the world"}
projects <- astronauts %>% 
  select(year_of_mission, 
         mission_title) %>%
  group_by(mission_title) %>%
  unique() %>%
  ungroup() 


yearly_projects <- projects %>% 
  mutate(mission_title = as_factor(mission_title)) %>% 
  group_by(year_of_mission) %>% 
  summarise(total = length(mission_title)) %>%
  ungroup()

projects_plot <- yearly_projects %>%
  ggplot(aes(year_of_mission, total)) +
  geom_point() +
  geom_line() + 
  xlab("Year") +
  ylab("Total") +
  ggtitle("Global Yearly Project Amount")

ggplotly(projects_plot)
```


Beside the number of projects, the mission amount which needed to be carried out by each shuttle launched could also be used for identifying the trends of astronautical exploration, because each mission in a project is for helping human to explore the universe. According to figure \@ref(fig:yearly-mission), it seems likely that there are four times of astronautical exploration rush in human history. In those period, humans set many missions to help the space exploration. The first rush happened in 1985. There were over 60 missions has been carried out in the space , and it was the peak for astronautical exploration. Human has started to show their enthusiasm for astronautical exploring. The duration and amount of mission has remained in a high level in the following thirty-year. Four years later after 1985, the second rush appeared and lasted thirteen years. In that period, average 40 missions has been carried out in universe per year. The third and forth peak also happened in the following years, which lasted three and two years, respectively. Although, the amount of missions decreased after 2009, the mission amount still remained in above 10 cases level per year until now. Overall, human never stop their steps to explore the space. astronautical exploration has experienced a increase trend from 1961 and remained in a high level for many years.

```{r yearly-mission, fig.cap= "Yearly astronautical exploration mission amount in the world (total mission in yearly projects)"}
yearly_mission <- astronauts %>% 
  group_by(year_of_mission) %>% 
  summarise(total_mission = length(total_number_of_missions))

yearly_mission %>% 
  ggplot(aes(year_of_mission, total_mission)) +
  geom_col(fill = "#6cc5eb") + 
  geom_line(lwd = 1, colour = "red") +
  geom_text(aes(label=year_of_mission, angle = 90), size = 3, hjust = -0.1, vjust = 0.3, position = "stack") + 
  ylim(0,70) + 
  xlab("Year") + 
  ylab("Total") + 
  ggtitle("Global Yearly Mission Amount")
```

In addition, an astronautical exploration race has launched along with the human desire for space. Figure \@ref(fig:countries-line-chart) and figure \@ref(fig:countries-bar-chart) indicate the astronautical mission amount trends of each country and global astronautical mission share of each country, respectively. According to figure \@ref(fig:countries-line-chart), the main forces of space exploration were United State and U.S.S.R, which were in the same level before 1981. After that, however, U.S. increased the mission amount significantly, and lead the world level over thirty-two years. On the other hand, based on figure \@ref(fig:countries-bar-chart), U.S. and U.S.S.R/Russia seem play an important role in astronautical exploration in the world. But U.S was still the leader in the astronautical field, and it has occupied the most astronautical exploration missions in the world. @blamont2005international explained that the awareness of universe, and technical advantage have contributed the leader level to United State. However, even though U.S. and U.S.S.R dominated the astronautical field from 1961 and 1977, there were some new countries join that field in 1978 based on figure \@ref(fig:countries-bar-chart). And after that the trend of new countries joining the universe exploration is increasing in general. In conclusion, although the technical advantages has increased the gaps of each country, and contributed the leader level to United State, many countries still join the astronautical exploration. And more and more countries would join in the exploration, which might become the trend.

```{r countries-line-chart, fig.cap="Yearly mission amount of each countries"}
countries <- astronauts %>%
  group_by(year_of_mission,nationality) %>% 
  summarise(total_mission = length(nationality)) %>%
  ungroup() %>% 
  mutate(nationality = ifelse(total_mission <4, "other", nationality))

countries_plot <- countries %>%
  ggplot(aes(year_of_mission, total_mission)) +
  geom_line(aes(colour = nationality)) + 
  labs(colour = "country") + 
  xlab("Year") + 
  ylab("Count") + 
  ggtitle("Country Mission Amount")

ggplotly(countries_plot)
```

```{r countries-bar-chart, fig.cap="The global percentage of each country's mission"}
countries_two <- astronauts %>%
  select(nationality, year_of_mission) %>%
  mutate(nationality = ifelse(nationality == "U.S." | nationality=="U.S.S.R/Russia", nationality, "other"))

countries_two %>% 
  ggplot() + 
  geom_bar(aes(x = year_of_mission, fill = nationality), position = "fill") + 
  scale_fill_manual(values = c("#eb3434", "#34bdeb", "#e2eb34")) + 
  xlab("Year") + 
  ylab("Count") + 
  theme_classic() + 
  guides(fill = guide_legend(title = "country")) + 
  ggtitle("Global Mission Percentage")
```

## The tech of astronautic exploration development

### Number of extravehicular (EVA) operations per year

This Figure \@ref(fig:plotEVA) describe the Number of extravehicular (EVA) operations per year. Add up all the EVA in the same year and make this bar chart. Since the first spacewalk performed by Alexei Leonov in 1965, the number of extravehicular activity (EVA) was performed steadily increased especially around year 2000 over than 300 times per year. After year 2000, the number of times of EVA has decreased, it does not mean that the development of spacefights is regressive. Figure \@ref(fig:plotmissionhours) shows the execution duration for each mission and the results is dramatically increased.  

```{r select-raw-data}
EVA_number_year <- astronauts %>%
  select(year_of_mission, eva_hrs_mission) %>%
  #arrange(year_of_mission)
  group_by(year_of_mission) %>%
  summarise(sum(eva_hrs_mission)) %>%
  rename(Total = `sum(eva_hrs_mission)`, Year = year_of_mission)
```

```{r plotEVA, fig.cap="Number of EVA operation per year"}
EVA_number_year %>%
  ggplot(aes(x = Year, y = Total)) +
  geom_bar(stat="identity") +
  theme(panel.grid.major =element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),axis.line = element_line(colour = "black")) +
  ylab("Number of EVA") +
  ggtitle("Number of Extravehicular (EVA) Operations per Year")
```


### Proportion of Space Travelers in Different Spaceflights

Based on the article written by Corlett,  Stavnichuk and Komarova in 2020, divided the duration of mission hours into three categories could observe the progress of space traveler more intuitively. These three categoris include pariticipant short spacelights less than 96 hours, intermediate spaceflights between 96 to 480 hours and long spaceflights more than 480 hours. When the spaceflights first developed, all the duration of mission is less than 96 hours. After that the time astronauts spent in space increased through the spaceflights histories. The recent trend is the time spent on each mission exceeds 480 hours which shows in Figure \@ref(fig:plotmissionhours). @NASA explained the potential reason related to the launch of the International Space Station (ISS) program in 1998. 

```{r total-astronausts}
total_astronauts <- astronauts %>%
  select(name, year_of_mission, hours_mission) %>%
  arrange(year_of_mission) %>%
  count(year_of_mission)
```  
  
```{r mission-less-96h}  
astronauts_96 <- astronauts %>%
  select(year_of_mission, hours_mission) %>%
  filter(hours_mission < 96 ) %>%
  count(year_of_mission)
```

```{r join-96-total}
join_96 <- full_join(astronauts_96,total_astronauts,  by = "year_of_mission") %>%
  arrange(year_of_mission)

join_96[is.na(join_96)] <- 0
prop_96 <- join_96 %>%
  rename("number96" = n.x, total = n.y) %>%
  mutate(proportion_96 = (number96 / total)) %>%
  select(year_of_mission, proportion_96)
```


```{r mission-between-96-480h}
astronauts_96_480<- astronauts %>%
  select(year_of_mission, hours_mission) %>%
  # mutate(total = sum(year_of_mission)) %>%
  filter(hours_mission > 96 & hours_mission < 480) %>%
  count(year_of_mission)
```

```{r join--96-480-total}
join_96_480 <- full_join(astronauts_96_480, total_astronauts, by = "year_of_mission") %>%
  arrange(year_of_mission)

join_96_480[is.na(join_96_480)] <- 0 
prop96_480 <- join_96_480 %>%
  rename("number96_480" = n.x, total = n.y) %>%
  mutate(proportion96_480 = (number96_480 / total)) %>%
  select(year_of_mission, proportion96_480)
```

```{r mission-over-480h}
astronauts_480 <- astronauts %>%
  select(year_of_mission, hours_mission) %>%
  # mutate(total = sum(year_of_mission)) %>%
  filter(hours_mission > 480) %>%
  count(year_of_mission)
```

```{r join-three-tables}
join_480 <- full_join(astronauts_480, total_astronauts, by = "year_of_mission") %>%
  arrange(year_of_mission)

join_480[is.na(join_480)] <- 0 
prop_480 <- join_480 %>%
  rename("number480" = n.x, total = n.y) %>%
  mutate(proportion_480 = (number480 / total)) %>%
  select(year_of_mission, proportion_480)
```

```{r plotmissionhours, fig.cap="Proportion of space travelers in different spaceflights"}
mission <- full_join(prop_96, prop96_480, by = "year_of_mission") %>%
  full_join(prop_480, by = "year_of_mission") %>%
  rename("< 96h" = proportion_96, "96 to 480h" = proportion96_480, "> 480h" = proportion_480)

mission1 <- mission %>%
  pivot_longer("< 96h":"> 480h", names_to = "categories", values_to = "proportion") %>%
  mutate(categories = as_factor(categories)) %>%
  rename("Categories" = "categories")
# mutate(categories = fct_relevel(categories, c("96 to 480h", "> 480h", "< 96h")))

ggplot(mission1, aes(year_of_mission, proportion, Categories, fill = Categories)) +
  geom_bar(stat="identity",position="fill") +
  scale_fill_manual("Categories", values = c("96 to 480h" = "#F5B041", "< 96h" = "#F5CBA7", "> 480h" = "#D35400")) +
  ggtitle("Proportion of Space Travelers that Participated in Short, Intermediate and Long Spaceflights")+
  theme(panel.grid.major=element_line(colour=NA)) +
  xlab("Year") +
  ylab("Proportion of Astronauts") +
  theme(text = element_text(size = 9), axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12))
```

## Women performance in astronautical exploration

According to figure \@ref(fig:gender), while women have also appeared in space exploration, men have been the main force in this 49-year period. However, we cannot ignore the contribution of female astronaut in space exploration. @jacksona2019women indicated that women has preformed essential work in astronautical area since 19th century. Furthermore, @jacksona2019women also illustrated that the first women travelled into space was in 1963, and the first American female astronaut appeared in 1983. Female astronauts also contribute to the exploration same as male astronauts. 

```{r gender, fig.cap="Astronaut gender amount comparison"}
gender <- astronauts %>%
  select(year_of_mission,sex) %>% 
  group_by(year_of_mission) %>% 
  count(sex) 


gender %>% 
  ggplot() + 
  geom_bar(aes(sex, fill = sex)) + 
  xlab("gender") + 
  ggtitle("Astronaut Gender Amount")
```

Female astronauts trend to increase in the space exploration. According to figure \@ref(fig:gender-percentage), although there was female astronauts appears in the space in 1963, there was a 20-year empty period for female astronauts until 1981. Such situation finished until 1982, and after that, female astronauts trend to increase in fluctuation along with year growing. And female astronauts have appeared in the astronautical exploration almost every year. 

```{r gender-percentage, fig.cap="Global gender share in astronautical exploration of each year "}
astronauts %>% 
  select(year_of_mission, sex) %>% 
  ggplot() + 
  geom_bar(aes(x = year_of_mission, fill = sex), position = "fill") + 
  xlab("year") + 
  ggtitle("Global Gender Share")
```

Overall, female astronauts have started to appear frequently in the space exploration, and experienced a general increase tendency. However, female astronauts are still account for a low share compared with male astronauts. While, @jacksona2019women explained that the possibilities and need for more females in astronautical exploration have been considered in the worldwide nowadays. Female astronauts would trend to contribute more in the space exploration in the future, and more and more female might appear in the universe.


# Conclusions 

In conclusion, Human shows their interesting and enthusiasm for universe. The astronautical exploration projects and missions experienced a increase trend since the first humans appeared in the universe. And there were a space exploration rush lasted almost 24 years in the world. In that period, the number of project and mission are the highest level from 1961 to today. Meanwhile, many countries also join in the project of the astronautical exploration, which might contribute the development of the space exploration. 

Furthermore, the technologies of the space exploration also have developed. Through the spaceflight activities of space travelers exploration, the overall EVA number steadily increased ten years before, but it decreased in recent years. However, it does not mean the development of the spaceflights is regressing. Since the launch of the ISS program in 1998 (@NASA), the duration of almost each mission task is more than 480 hours.

In addition, female astronauts also appear in the history of astronautical exploration. Although male astronauts are still the main force in the universe, the number of female astronauts trend to increase from 20th century. And the appear of female astronauts might change the history of space exploration and help human with the exploration.


# Acknowledgement

The programming language used to analyse the trends of global astronautical exploration is R (4.0.2) [@R]. And the platform which used for running R language is Rstudio (1.3.1056) [@Rstudio].

Following packages has been included in our Rmd file.

- package tidyverse (1.3.0) [@tidyverse],
- package dplyr (1.0.1) [@dplyr],
- package plotly (4.9.2.1) [@plotly],
- package ggplot2 (3.3.2) [@ggplot2]


# References
